Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-11 podc16/paper2.tex
work on describing the linearization algorithm + wrote an introductory paragraph to describe the main idea.

%\todo{Let's base everything on C-Structs; it will be easier. We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and it produces responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. A c-struct set is a set, containing the element $\bot$, with an append operator
$\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct). In the rest of the paper, unless otherwise specified, the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.

In the next sections we present our specification of leaderless generalized consensus algorithms, which is parameterized over a \textit{dependency-set} algorithm (\Cref{dep-algo}) and a \textit{map-agreement} algorithm (\Cref{map-algo}). 
%Processes proposals of commands to the \textit{Generalized Consensus} layer
%We also consider a deterministic state-machine $M$ whose set of commands is $C$. 
%We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect on the state machine as executing $c_b$ before $c_a$. We assume that a command does not commute with itself.
%We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\notin R$ then $c_a$ and $c_b$ commute.

@@

The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$). For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
Given a graph $g$, let $\opn{sccs}\left( g \right)$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $scc_1$ to $scc_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( scc_1 \right)$, $v_2 \in V\left( scc_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$. Note that $\opn{sccs}\left( g \right)$ is necessarily a forest.

@@

>>>>

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and it produces responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. A c-struct set is a set, containing the element $\bot$, with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. 
We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$. 
C-structs are partially ordered by the relation $\preceq$ defined such that $\sigma_1\preceq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct). 
In the rest of the paper the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
\todo{Present generalized consensus here?}

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on. Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable. 
Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{graph-proc}.
Any other process can run the three phases independently of $p$ and eventually execute the command.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when at least a majority of the processes are correct.

%In the next sections we present our specification of leaderless generalized consensus algorithms, which is parameterized over a \textit{dependency-set} algorithm (\Cref{dep-algo}) and a \textit{map-agreement} algorithm (\Cref{map-algo}).

@@

\subsection{Graph Processing}
\label{graph-proc}

The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$). For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.

Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$. Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order. For example, if $V\left( g
\right)=\aset{c_1,c_2,c_3,c_4}$
and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.

A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
Let $\opn{TOs}\left( g \right)$ be the set of total orders $\leq$ such that $\preceq_g \subseteq \leq$

Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.

When $\hat{o}$ is a total order in $\opn{TOs}\left( \hat{g} \right)$, define the total order $\leq_{\hat{o}}^g$ on $V\left( g \right)$ such that $v_1 \leq_{\hat{o}}^g v_2$ iff (1) there exists $V_1,V_2\in V\left(\hat{g}\right)$ such that $v_1\in V_1$, $v_2\in V_2$, and $\left( V_1,V_2 \right)\in \hat{o}$, or (2) there exists $V\in V\left(\hat{g}\right)$ such that $v_1,v_2\in V\left( V \right)$ and $\left( v_1,v_2 \right)\in f\left[ V\left( V \right) \right]$.
We say that $\leq_{\hat{o}}^g$ is a \emph{linearization} of $g$.

Property: if $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, $E\left( g_1 \right)\subseteq E\left( g_2 \right)$, $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$, then $\bot\bullet l1 \preceq \bot\bullet l2$.

\begin{comment}
Define $\opn{Lin}_g\left( \leq \right) = f\left[ V_1 \right] \cdot F\left[ V_2 \right] \cdot F\left[ V_n \right]$, where $V_i\in V\left( \hat{g} \right)$

@@

\end{comment}
